---
title: "p8105_hw1_sl4659"
author: "Shenglin Liu"
date: "9/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r dataframe}
library(tidyverse)
test_df = tibble(
  # a random sample of size 8 from a standard Normal distribution
  vec_nor = rnorm(8),
  # logical vector indicating whether elements of the sample are greater than 0
  vec_log = vec_nor > 0,
  # character vector of length 8
  vec_char = c("My","name","is","Shenglin","Liu","I","Love","R"),
  # factor vector of length 8, with 3 different levels
  vec_fac = factor(c("high","low","low","medium","high","medium","low","low"))
)
# take the mean of each variable
mean(pull(test_df, vec_nor))
mean(pull(test_df, vec_log))
mean(pull(test_df, vec_char))
mean(pull(test_df, vec_fac))
```
It works for vec_nor and vec_log and doesn't work for vec_char and vec_fac.

```{r numeric, message = FALSE, warning = FALSE, results = "hide"}
# as.numeric function
as.numeric(pull(test_df, vec_nor))
as.numeric(pull(test_df, vec_log))
as.numeric(pull(test_df, vec_char))
as.numeric(pull(test_df, vec_fac))
```
All variables can be converted to numeric except for vec_char.
It helps explaining why mean() doesn't work for vec_char. However, mean() doesn't work for vec_fac even if it can be converted to numeric.

```{r second}
# convert logical to numeric, and multiply the random sample by the result
as.numeric(pull(test_df, vec_log))*pull(test_df, vec_nor)
# convert logical to factor, and multiply the random sample by the result
as.factor(pull(test_df, vec_log))*pull(test_df, vec_nor)
# convert logical to factor, and then convert the result to numeric, and multiply the random sample by the result
as.numeric(as.factor(pull(test_df, vec_log)))*pull(test_df, vec_nor)
```
## Problem 2

```{r dataframe2}
test2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_log2 = x+y > 1,
  vec_num2 = as.numeric(vec_log2),
  vec_fac2 = as.factor(vec_log2)
)
```
The size of the dataset is `r nrow(test2_df)*ncol(test2_df)`.
The mean of x is `r mean(pull(test2_df, x))`.
The median of x is `r median(pull(test2_df, x))`.
The standard deviation of x is `r sd(pull(test2_df,x))`.
The proportion of cases for x + y > 1 is `r sum(pull(test2_df,vec_log2), na.rm = TRUE)/500`.


